Blind Search for 21-cm Absorption Systems in New Generation Chinese Radio Telescopes

### 2016-12-23 ms_21cm.tex
update numbers

\abstract{
Neutral hydrogen clouds are known to exist in the Universe, however their spatial
distributions and physical properties are poorly understood. Such missing information
can be studied by the new generation Chinese radio telescopes through a blind searching
of 21-cm absorption systems. We forecast the capabilities of surveys of 21-cm absorption
systems by two representative radio telescopes in China --
Five-hundred-meter Aperture Spherical radio Telescope (FAST) and
Tianlai 21-cm cosmology experiment (Tianlai).
Facilitated by either the high sensitivity (FAST)
or the wide field of view (Tianlai) of these telescopes,
\tcr{???} 21-cm absorption systems can be discovered in a few years,
representing orders of magnitude improvement over the cumulative discoveries
in the past half a century. 
\keywords{}
}

@@

NVSS contains 2 million sources (with declination $\delta >-40^\circ$,
cover 82\% of the sky) stronger than 2.5
mJy at $\nu=$1.4 GHz \citep{1998AJ....115.1693C} and their flux
distribution $n_R(F)$ is given by \cite{1984ApJ...287..461C}.
Redshifted {\HI} clouds absorb lower frequency bands where
the radio sources are typically brighter by $\nu^{-0.7}$
\citep{1998AJ....115.1693C}, or $(1+z_{\rm \HI})^{0.7}$.
The error of the measurement $\Delta F$ is given by \tcr{[reference]}
\begin{equation}\label{eq.delta_F}
    \Delta F={\rm SEFD}/\sqrt{n_{\rm pol}\Delta\nu\Delta t},
\end{equation}
where SEFD \tcr{(or $T_{\rm sys}$ TBD)} is the system equivalent flux density, $n_{\rm pol}$
is the number of polarizations, $\Delta\nu$ is the line width and
$\Delta t$ is the integration time. $\Delta\nu=(u_{\rm width}/c)\nu_{\rm obs}$,
where $u_{\rm width}$ is the equivalent line width of the {\HI}
absorber. The properties of 21-cm absorbers are primarily derived from
followup studies of optical absorbers. The discovery rate from the
cross correlation is a lower bound on the expect number of absorbers,
since high column density systems in the CNM may systematically
obscure potential background optical sources \citep{2014PhRvL.113d1303Y}. There are
only 3 blind radio detections, and a survey may discover more systems
which are optically obscured. They would likely be cold and at high
column density. For sensitivity purposes, we treat all sources as
$u_{\rm width}=2\ {\rm km\ s}^{-1}$
\citep{1982ApJ...259..495W,2005ARA&A..43..861W}.
$\nu_{\rm obs}$ is the observation frequency. For redshifted
21-cm absorption systems, $\nu_{\rm obs}=1420\ {\rm MHz}/(1+z_{\rm \HI})$.
The integration time $\Delta t=(n_{\rm obs}\times 24\ {\rm h})\tau$,
where $n_{\rm obs}$ is the number of days an object is scanned,
and $\tau=\lambda_{\rm obs}/2\pi D\cos\delta$ is the fractional time the object
transits the FoV ($\lambda_{\rm obs}\ll D$ or $\delta\nrightarrow\pi/2$).
For 21-cm absorption systems $\lambda_{\rm obs}=21\ {\rm cm}\times (1+z_{\rm obs})$.
$\Delta F$ is redshift- (frequency-) independent, as $(1+z_{\rm obs})$ terms
from integration time and line width cancel out
in the equation (\ref{eq.delta_F}).

@@

FAST has a fixed 500-meter-dish, and a circle of 300 meter in diameter is held
in the correct parabolic shape, being used at any one time. This
enables its zenith angle up to $40^\circ$ and it can scan objects
in declination range $-14^\circ 12'<\delta<65^\circ 48'$.
The 19-beam feed-horn array at L band will be the primary survey
instrument for FAST. For a single day it can scan total of $0.762^\circ$
in declination $\delta$. FAST survey strategy scans each object
only once ($n_{\rm obs}=1$), and with its high sensitivity,
even faint sources are identified in a single day. 
We assume a one-month survey by FAST
around the celestial equator for simplicity. FAST observes
1.2 to 1.4 GHz (redshift $z_{\rm \HI}<0.2$), so objects on celestial equator
have average transit time $\Delta t \simeq 10\ {\rm sec.}$
Taking $\Delta\nu\simeq$ 9 kHz, $n_{\rm pol}=2$
(dual polarized system), and
${\rm SEFD}=1$ Jy for FAST.
we have $\Delta F \simeq\ 2.3 {\rm mJy}$.
Taking into account the $\nu^{-0.7}$ flux boost at lower
frequencies, $rF>10\Delta F$ requires $F\gtrsim 0.11\ {\rm Jy}$ on 1.4 GHz.

One-month-scan ($n_{\rm obs}=30$) around celestial equator by FAST will have
$N_R\simeq 10^4$ sources whose $F\gtrsim 0.11\ {\rm Jy}$ on 1.4 GHz.
We further assume the independency between $n_R(F)$, $n_R(z)$
and $n_{\rm\HI}(z)$, from $n_{\rm system}(z)$ by integrating
equation (\ref{eq.n_system}), there will be $N_{\rm system}\simeq 90 {\rm /month}$ absorption
systems found in redshift range $0<z<0.2$.

@@

For Tianlai, equation (\ref{eq.N_R_eff}) gives $N_R\simeq 1.3\times 10^4$,
and applying it to equation (\ref{eq.n_system}) again
we get $N_{\rm system}\simeq 80 {\rm /year}$ absorption
systems found in redshift range $0.58<z<0.78$.

\section{Conclusion}\label{sec.conclusion}
We forecast the capability of FAST and Tianlai telescopes to search 21-cm
absorption systems. According to our assumptions, FAST is able
to find $\simeq 90$ systems per month
whereas TianLai can find $\simeq 80$ systems per year.
Comparing the results between two quite different telescopes --
high sensitivity FAST and large FoV Tianlai, we find the former is
more efficient in looking for fainter radio sources
($n_R(F)\sim F^{-2.5}$) for possible foreground absorptions.
Future improvements in the sensitivity on Tianlai enable it
more effectively looking for highest signal-to-noise systems
over a wider sky.

>>>>

\abstract{
Neutral hydrogen clouds are known to exist in the Universe, however their spatial
distributions and physical properties are poorly understood. Such missing information
can be studied by the new generation Chinese radio telescopes through a blind searching
of 21-cm absorption systems. We forecast the capabilities of surveys of 21-cm absorption
systems by two representative radio telescopes in China --
Five-hundred-meter Aperture Spherical radio Telescope (FAST) and
Tianlai 21-cm cosmology experiment (Tianlai).
Facilitated by either the high sensitivity (FAST)
or the wide field of view (Tianlai) of these telescopes,
\tcr{??? [more than a thousand for FAST]} 21-cm absorption systems can be discovered in a few years,
representing orders of magnitude improvement over the cumulative discoveries
in the past half a century. 
\keywords{}
}

@@

NVSS contains 2 million sources (with declination $\delta >-40^\circ$,
cover 82\% of the sky) stronger than 2.5
mJy at $\nu=$1.4 GHz \citep{1998AJ....115.1693C} and their flux
distribution $n_R(F)$ is given by \cite{1984ApJ...287..461C}.
Redshifted {\HI} clouds absorb lower frequency bands where
the radio sources are typically brighter by $\nu^{-0.7}$
\citep{1998AJ....115.1693C}, or $(1+z_{\rm \HI})^{0.7}$.
The error of the measurement $\Delta F$ is given by\footnote{https://science.nrao.edu/facilities/vla/docs/manuals/oss/performance/sensitivity} \tcr{[confirm this reference]}
\begin{equation}\label{eq.delta_F}
    \Delta F={\rm SEFD}/\sqrt{n_{\rm pol}\Delta\nu\Delta t},
\end{equation}
where SEFD \tcr{(or $T_{\rm sys}$ TBD)} is the system equivalent flux density, $n_{\rm pol}$
is the number of polarizations, $\Delta\nu$ is the line width and
$\Delta t$ is the integration time. $\Delta\nu=(u_{\rm width}/c)\nu_{\rm obs}$,
where $u_{\rm width}$ is the equivalent line width of the {\HI}
absorber. The properties of 21-cm absorbers are primarily derived from
followup studies of optical absorbers. The discovery rate from the
cross correlation is a lower bound on the expect number of absorbers,
since high column density systems in the CNM may systematically
obscure potential background optical sources \citep{2014PhRvL.113d1303Y}. There are
only 3 blind radio detections, and a survey may discover more systems
which are optically obscured. They would likely be cold and at high
column density. For sensitivity purposes, we treat all sources as
$u_{\rm width}=2\ {\rm km\ s}^{-1}$
\citep{1982ApJ...259..495W,2005ARA&A..43..861W}.
$\nu_{\rm obs}$ is the observation frequency. For redshifted
21-cm absorption systems, $\nu_{\rm obs}=1420\ {\rm MHz}/(1+z_{\rm \HI})$.
The integration time $\Delta t=(n_{\rm obs}\times 24\ {\rm h})\tau$,
where $n_{\rm obs}$ is the number of days an object is scanned,
and $\tau=\lambda_{\rm obs}/2\pi D\cos\delta$ is the fractional time the object
transits the FoV ($\lambda_{\rm obs}\ll D$ or $\delta\nrightarrow\pi/2$).
For 21-cm absorption systems $\lambda_{\rm obs}=21\ {\rm cm}\times (1+z_{\rm obs})$.
$\Delta F$ is redshift- (frequency-) independent, as $(1+z_{\rm obs})$ terms
from integration time and line width cancel out
in the equation (\ref{eq.delta_F}).

@@

FAST has a primary dish of 500 meter in diameter. The effective
antenna aperture of FAST is 300 m, in diameter for any zenith angle
up to $26^\circ$ and decreases to 200 m for a maximum zenith angle 
of $40^\circ$ \citep{2016RaSc...51.1060L}.
%This enables its zenith angle up to $40^\circ$ and it can scan objects
%in declination range $-14^\circ 12'<\delta<65^\circ 48'$.
The 19-beam feed-horn array at L band will be the primary survey
instrument for FAST. For a single day it can scan total of \tcr{$0.5^\circ$ (TBD)}
in declination $\delta$. FAST survey strategy scans each object
only once ($n_{\rm obs}=1$), and with its high sensitivity,
even faint sources are identified in a single day. 
We assume a one-month survey by FAST
around the celestial equator for simplicity. FAST observes
1.02 to 1.42 GHz (redshift $z_{\rm \HI}<0.39$), so objects on celestial equator
have average transit time $\Delta t \simeq \tcr{12}\ {\rm sec.}$
Taking $\Delta\nu\simeq$ 9 kHz, $n_{\rm pol}=2$
(dual polarized system), and
${\rm SEFD}=1$ Jy for FAST.
we have $\Delta F \simeq\ 2.3\ {\rm mJy}$.
Taking into account the $\nu^{-0.7}$ flux boost at lower
frequencies, $rF>10\Delta F$ requires $F\gtrsim 0.11\ {\rm Jy}$ on 1.4 GHz.

One-month-scan ($n_{\rm obs}=30$) around celestial equator by FAST will have
$N_R\simeq 10^4$ sources whose $F\gtrsim 0.11\ {\rm Jy}$ on 1.4 GHz.
We further assume the independency between $n_R(F)$, $n_R(z)$
and $n_{\rm\HI}(z)$, from $n_{\rm system}(z)$ by integrating
equation (\ref{eq.n_system}), there will be $N_{\rm system}\simeq \tcr{210} {\rm /month}$ absorption
systems found in redshift range $0<z<0.39$.

@@

For Tianlai, equation (\ref{eq.N_R_eff}) gives $N_R\simeq 1.3\times 10^4$,
and applying it to equation (\ref{eq.n_system}) again
we get $N_{\rm system}\simeq \tcr{80} {\rm /year}$ absorption
systems found in redshift range $0.58<z<0.78$.

\section{Conclusion}\label{sec.conclusion}
We forecast the capability of FAST and Tianlai telescopes to search 21-cm
absorption systems. According to our assumptions, FAST is able
to find $\simeq 210$ systems per month
whereas TianLai can find $\simeq 80$ systems per year.
Comparing the results between two quite different telescopes --
high sensitivity FAST and large FoV Tianlai, we find the former is
more efficient in looking for fainter radio sources
($n_R(F)\sim F^{-2.5}$) for possible foreground absorptions.
Future improvements in the sensitivity on Tianlai enable it
more effectively looking for highest signal-to-noise systems
over a wider sky.
